---
title: "How to use eBird filter compare"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ebirdfiltercompare}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## First time setup

Install R and RStudio and create a new Rstudio project, as needed. See README for details.

## Load the R package

```{r setup}
library(ebirdfiltercompare)
```

## Obtain the filter html files that you want to use

Skip this entire step if you just want to use the built-in example files.

1. Save a copy of each your current eBird filters
  - Using Google Chrome or Mozilla Firefox, log in to eBird Admin
    - Note: Google Chrome browser seems to be saving html files in different way, so we recommend using Firefox to save html files
  - Go to the filters tab
  - Right-click on filter page > "View Page Source"
  - File > Save Page As
    - We recommend saving to a `filter_htm` folder inside your project directory
  - Select 'Webpage, HTML Only'
2. After you have done this, take a second to double check you have the right number of files saved in this folder -- you should have as many files as you have filters for your region or else it will cause trouble!
3. Double check that there are no spaces or strange characters in your filter filenames

## Specify portions of filter names to be deleted

These can be R regular expressions (see `?base::regex` for details). This example removes the word Wisconsin, two hyphens, two digits, and a period from the beginning of filter names. For custom usage, we recommend starting with an empty string (like in this `filter_suffix` example) for both, and adjusting regular expressions needed on a second run.

```{r prefix_suffix}
filter_prefix <- 'Wisconsin--\\d{2}\\.'
filter_suffix <- ''
```

## Get list of filter filenames

Scan a directory for the filter html files you downloaded. Running it with no arguments, like below, uses built-in example files. To use your files, provide the path to the folder you want to use. For example, you would use `get_filenames(path = 'filter_htm')`, if they are located in the `filter_htm` folder at the top level of your R project directory.

```{r get_filenames}
get_filenames() -> files
```

## Export CSV list of regions for manual ordering

The files will initially be in alphabetical order, and the resulting regions here will be in that same order. This step exports a CSV you can manually edit to change the order of the regions as needed for eventual display. By default, the function outputs to the `output` folder in the current working directory and creates that folder if needed.

```{r export_region_list_for_ordering}
export_region_list_for_ordering(files, filter_prefix, filter_suffix) -> regions
```

After you run this function, check your output directory for `regions.csv`, edit the file (e.g., in Excel) to adjust the order, and save the new file as `ordered_regions.csv`.

## Import manually ordered list of regions

This step re-imports your ordered regions. Calling the function with no argument, like below, uses a built-in example file. For custom usage, you'll want to provide the file argument to the location of your own file, e.g. `import_ordered_region_list(file_path('output', 'ordered_regions.csv'))`

```{r import_ordered_region_list}
import_ordered_region_list() -> ordered_regions
```

## Check ordered_regions vs. regions for consistency

```{r check_regions}
check_regions(regions, ordered_regions)
```

## Collect the taxa that appear in your filters

```{r compile_taxonomy}
compile_taxonomy(files, regions) -> species
```

## Extract filter data from the filter HTML files

This can currently take about 30 seconds per filter due to many nested for loops in the code. This may be further optimized in the future for faster runs.

```{r crunch_filters}
crunch_filters(files, species, ordered_regions, filter_prefix, filter_suffix) -> data
```

## Generate a PDF showing filter comparisons across regions

By default, the function outputs to the `output` folder in the current working directory and creates that folder if needed.

```{r generate_pdf}
generate_pdf(data, species, ordered_regions)
```

## Generate a CSV index to page numbers in the PDF

By default, the function outputs to the `output` folder in the current working directory and creates that folder if needed.

```{r generate_index}
generate_index(species)
```
